---
title: "Final project; Data processing - Data Science for Linguists (LING 2340)
Fall 2022"
subtitle: "Data Processes 3"
author: "Gianina Morales"
date: "12/15/2022"
output:
  github_document:
    toc: TRUE
always_allow_html: true
---
# Data processing - Part 3 (final report)

---

```{r setup, message=FALSE, warning=FALSE}
library(tidytext)
library(tidyverse)
library(tm)
library(stopwords)
library(stringi)
library(topicmodels)
library(lda)
library(stm)
library(ldatuning)
library(rmarkdown)
library(kableExtra)
knitr::opts_chunk$set(fig.fullwidth=TRUE, fig.path = "../Images/Data_processes3/", cache=TRUE)
```

# Trends over decades

## Applying a new method: Structured Topic Models (STM)



**Preparing the data**

```{r}
#Loading Rds with all the corpus
total_corpus <- read_rds("../Private/master_tidy_corpus.Rds")

# preparing data frame for topic modeling
total_corpus_sparse <- total_corpus %>% 
  mutate(document= paste(decade, id, sep= "_")) %>% 
# creating DTM
  count(document, word) %>% 
  cast_sparse(document,word,n)
```

### Topic model

```{r Ranking_of_STM_topics}
set.seed(123)
topic_model_corpus <- stm(total_corpus_sparse, K=25)
#Summary
summary(topic_model_corpus)
#Ranking of topics
topics_rank <- plot(topic_model_corpus, "summary", n=5, text.cex= 0.6, main= "Topic shares of the entire corpus", xlab= "estimated share of topics")
  
```

### Results 

**1. Beta metric**

Top probability words by topic

```{r Beta_matrix}
#topic- word probability
betaSTM <- tidy(topic_model_corpus, matrix ="beta") %>% 
  group_by(topic) %>% 
  slice_max(beta, n=50) %>% 
  arrange(-beta) %>% 
  mutate(rank= row_number()) %>% 
  ungroup() %>% 
  select(-beta) %>%
  arrange(topic) %>% 
  pivot_wider(names_from = "topic", 
              names_glue = "topic {.name}",
              values_from = "term") %>% 
    select(-rank)
knitr::kable(betaSTM , options = list(rows.print=20))
```

**2. FREX**

FREX is a matrix that presents a balance between words that are frequent in probability but exclusive to the topic of interest.

```{r Frex_matrix, message=FALSE, warning=FALSE}
#FREX metric
devtools::install_github("juliasilge/tidytext")
frexSTM <- tidy(topic_model_corpus, matrix= "frex")%>% 
  group_by(topic) %>% 
  slice_head(n=20) %>% 
  mutate(rank= row_number()) %>% 
  ungroup() %>% 
  pivot_wider(names_from = "topic", 
              names_glue = "topic {.name}",
              values_from = "term") %>% 
    select(-rank)
knitr::kable(frexSTM, options = list(rows.print=20))
```

Useful only in some cases. Sometimes the words are related to authors.

I applied also the Lift metric (words calculated by dividing the topic-word distribution--beta-- by the word count probability distribution in each topic), but the result were not very useful because of the amount of words from other languages and names without format included.

### Analysis

**1. Topics named**

Considering the words in each metric, I analyzed the topics qualitatively, comparing the results from STM with my results by decade with the LDA method. My identification of topics is in the table.

```{r STM_topics_identified_table}
#Loading Rds with my qualitative identification of STM topics
STM_topics <- read_csv("../Private/all_topics.csv")
STM_topics_table <- STM_topics %>% 
  select(c("Topic_names", "STM_topic")) %>% 
  knitr::kable(booktabs = T, col.names = c("Topic names", "STM topic"), align = c("l", "l")) %>% 
  kable_classic (full_width = F)
STM_topics_table
STM_topics_table %>% 
    save_kable("../Images/Data_processes3/topicsSTM.png")
```

**2. Topics by decade**

I applied the gamma distribution to observe the decade per topic probability.

```{r Probability_topics_1969-1979, message=FALSE, warning=FALSE}
#gamma matrix
alldecades_gamma <- tidy(topic_model_corpus, matrix="gamma", document_names=rownames(total_corpus_sparse)) %>% 
  #separation of document name between decade and article id
   separate(document, c("decade", "article"), sep = "_", convert = TRUE)
#visualization decade 1
alldecades_gamma %>%
  filter(decade == "decade1") %>% 
  mutate(topic=factor(topic)) %>% 
  ggplot(aes(topic, gamma, fill=topic))+
  geom_boxplot(alpha=0.7, show.legend=F)+
  facet_wrap(vars(decade), ncol=1)+
  labs(title = "Probability of topics per decade, 1969 to 1979", y = expression(gamma))
```

Each dot is a document marking the document-topic probabilities in the decade.

```{r Probability_topics_1980-1989}
#visualization decade 2
alldecades_gamma %>%
  filter(decade == "decade2") %>% 
  mutate(topic=factor(topic)) %>% 
  ggplot(aes(topic, gamma, fill=topic))+
  geom_boxplot(alpha=0.7, show.legend=F)+
  facet_wrap(vars(decade), ncol=1)+
  labs(title = "Probability of topics per decade, 1980 to 1989", y = expression(gamma))
```

Each dot is a document marking the document-topic probabilities in the decade.

```{r Probability_topics_1990-1999}
#visualization decade 3
alldecades_gamma %>%
  filter(decade == "decade3") %>% 
  mutate(topic=factor(topic)) %>% 
  ggplot(aes(topic, gamma, fill=topic))+
  geom_boxplot(alpha=0.7, show.legend=F)+
  facet_wrap(vars(decade), ncol=1)+
  labs(title = "Probability of topics per decade, 1990 to 1999", y = expression(gamma))
```

Each dot is a document marking the document-topic probabilities in the decade.

```{r Probability_topics_2000-2009}
#visualization decade 4
alldecades_gamma %>%
  filter(decade == "decade4") %>% 
  mutate(topic=factor(topic)) %>% 
  ggplot(aes(topic, gamma, fill=topic))+
  geom_boxplot(alpha=0.7, show.legend=F)+
  facet_wrap(vars(decade), ncol=1)+
  labs(title = "Probability of topics per decade, 2000 to 2009", y = expression(gamma))
```

Each dot is a document marking the document-topic probabilities in the decade.

```{r Probability_topics_2010-2019}
#visualization decade 5
alldecades_gamma %>%
  filter(decade == "decade5") %>% 
  mutate(topic=factor(topic)) %>% 
  ggplot(aes(topic, gamma, fill=topic))+
  geom_boxplot(alpha=0.7, show.legend=F)+
  facet_wrap(vars(decade), ncol=1)+
  labs(title = "Probability of topics per decade, 2010 to 2019", y = expression(gamma))
```

Each dot is a document marking the document-topic probabilities in the decade.

```{r Probability_topics_2020-2022}
#visualization decade 6
alldecades_gamma %>%
  filter(decade == "decade6") %>% 
  mutate(topic=factor(topic)) %>% 
  ggplot(aes(topic, gamma, fill=topic))+
  geom_boxplot(alpha=0.7, show.legend=F)+
  facet_wrap(vars(decade), ncol=1)+
  labs(title = "Probability of topics per decade, 2020 to 2022", y = expression(gamma))
```

Only two and a half year.That explain the difference in some topics.

**3. Regression**

```{r estimate_effect}
#Documents
effects <-
  estimateEffect(
    1:25 ~ as.factor(decade),
    topic_model_corpus,
    total_corpus %>% distinct(id, decade)%>% arrange(id)
  )
summary(effects)
tidy(effects)
```

```{r}
sessionInfo()
```

